WDIAS : A Microservices-Based Weather Data Integration and Assimilation System\\

### 2021-01-01 Thesis.tex
Added feature comparison of the systems
After Dr. Dilum reviewed the Viva thesis

\usepackage{setspace}
\usepackage{tabu}
\usepackage[acronym,nopostdot,nonumberlist]{glossaries}
\makeglossaries
\usepackage{import}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{dirtytalk}
\usepackage{tabulary}
\usepackage[style=ieee]{biblatex}
\usepackage[titletoc]{appendix}
\usepackage{siunitx}

@@

        %\vfill
       \vspace{72pt}
        \large
        Department of Computer Science and Engineering\\
        \vspace{24pt}
        University of Moratuwa\\
        Sri Lanka\\
        \vspace{32pt}
        August 2020

@@

\acrfull{nwm} utilize data collected via diverse sources such as automated weather stations, radars, air balloons, and satellite images. Before using such multimodal data in a \acrshort{nwm}, it is necessary to transcode data into a format that can be ingested by the \acrshort{nwm}. Moreover, the data integration system's response time needs to be relatively low to forecast and monitor time-sensitive weather events like hurricanes, storms, and flash floods that require rapid and frequent execution of \acrshort{nwm}. The resulting weather data also need to be accessed by many researchers and third-party applications such as logistic and agricultural insurance firms.
Existing weather data integration systems are based on monolithic or client-server architectures; hence, unable to benefit from novel computational models such as cloud computing and containerized applications. Moreover, most of these software are proprietary or closed-source, making it difficult to customize them for an island like Sri Lanka with different weather seasons.
Therefore, in this research, we propose \acrfull{wdias} that utilizes modern architectural patterns such as microservices to achieve scalability, high availability, and low-cost operation based on cloud computing. The use of stateless microservices also enables \acrshort{wdias} to add new features on the fly with rollover capabilities. Moreover, \acrshort{wdias} provides a modular framework to integrate data from different sources, export into different formats, and add new functionality by adding extension modules.
We demonstrate the utility of \acrshort{wdias} using cloud-based experimental setup and weather-related synthetic workloads.

>>>>

\usepackage{setspace}
\usepackage{tabu}
\usepackage[acronym,nopostdot,nonumberlist]{glossaries}
\makeglossaries
\usepackage{import}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{dirtytalk}
\usepackage{tabulary}
\usepackage{multirow}
\usepackage[style=ieee]{biblatex}
\usepackage[titletoc]{appendix}
\usepackage{siunitx}

@@

        %\vfill
       \vspace{72pt}
        \large
        Department of Computer Science and Engineering\\
        \vspace{24pt}
        University of Moratuwa\\
        Sri Lanka\\
        \vspace{32pt}
        January 2021

@@

\acrfull{nwms} utilize data collected via diverse sources such as automated weather stations, radars, air balloons, and satellite images. Before using such multimodal data in a \acrshort{nwm}, it is necessary to transcode data into a format  %that can be 
ingested by the \acrshort{nwm}. Moreover, the data integration system's response time needs to be relatively low to forecast and monitor time-sensitive weather events like hurricanes, storms, and flash floods that require rapid and frequent execution of \acrshort{nwms}. The resulting weather data also need to be accessed by many researchers and third-party applications such as logistic and agricultural insurance firms.
Existing weather data integration systems are based on monolithic or client-server architectures; hence, unable to benefit from novel computational models such as cloud computing and containerized applications. Moreover, most of these softwares are proprietary or closed-source, making it difficult to customize them for an island like Sri Lanka with different weather seasons.
Therefore, in this research, we propose \acrfull{wdias} that utilizes %modern architectural patterns such as 
microservices to achieve scalability, high availability, and low-cost operation based on cloud computing. The use of stateless microservices also enables \acrshort{wdias} to add new features on the fly with rollover capabilities. Moreover, \acrshort{wdias} provides a modular framework to integrate data from different sources, export into different formats, and add new functionality by adding extension modules.
We demonstrate the utility of \acrshort{wdias} using a cloud-based experimental setup and weather-related synthetic workloads.

### 2021-01-01 introduction/sec_1-motivation.tex
After Dr. Dilum reviewed the Viva thesis

Weather forecast is essential to reduce the impact caused by natural disasters and to effectively manage natural resources like water.
The weather forecasts are performed using \acrfull{nwm} that use a set of configuration parameters and the current status of the environment such as temperature, rainfall, and wind direction as the inputs. For example, the \acrfull{wrf} \cite{MesoscaleMicroscaleMeteorologyLaboratoryWeatherModel} model is used to predict the rainfall, temperature, and wind direction. Typically multiple \acrshort{nwm} are executed on the same data with different configuration parameters. Then using the real-time observations, the most fitting model is selected. After removing the bias, the data extracted from a \acrshort{nwm} are then fed into another set of models/simulators. For example, FLO2D, a hydrologic model that predicts surface water level, could use precipitation estimates from \acrshort{wrf} to forecast flood levels.

\acrfull{curw} \cite{CUrWSL2017Lanka} is tasked with the mission to reduce water-related natural disasters in Sri Lanka. \acrshort{curw} \cite{CUrWSLObservedSL} is implementing own system with the idea of long-term maintainability rather than depending on existing weather data integration systems. \gkc{Because of existing weather systems are not open source to configure as required, and not flexible enough to change without much help from the maintainers.} \acrshort{curw} collects real-time data from weather stations and other relevant entities and generates daily forecasts. Then the generated results are shared with relevant public departments responsible for disaster planning, risks mitigation, and disaster management. One of the future goals of \acrshort{curw} is to open up data to the public enabling both societal and commercial use cases.

@@

\acrfull{dias} \cite{Kawasaki2018DataReduction} and \acrfull{madis} \cite{Macdermaid2005ArchitectureP2.39} are two of the well-known Weather Data Integration and Assimilation (WDIA) systems. However, those are proprietary systems. Further, while \acrfull{fews} \cite{Werner2013TheSystem} is free to use, but it is not open source. While a few other data integration systems exists, most of them are also proprietary or closed source. Thus, it is difficult and costly to customize such systems for an island like Sri Lanka which experience different kinds of weather seasons over the year \cite{NaveendrakumarFiveLanka}.
For example, \acrshort{fews} does not support FLO2D model integration, which is used by \acrshort{curw} to forecast water level.

\gc{To address such issues, there is a parallel researches also try to implement an open-data open-model framework called as Meta Scientific Modeling (MSM).} But the existing WDIA systems and emerging systems are also based on the monolithic or client service architecture; hence, cannot benefit from technologies such as cloud computing to achieve high scalability and availability. Moreover, such systems cannot be updated or reconfigured without taking the system off line. Furthermore, these systems are platform specific. For example, while \acrshort{fews} can run on either Windows or Linux, downtime is needed to introduce any changes to the system.

>>>>

Weather forecast is essential to reduce the impact caused by natural disasters and to effectively manage natural resources like water.
The weather forecasts are performed using \acrfull{nwms} that use a set of configuration parameters and the current status of the environment such as temperature, rainfall, and wind direction as the inputs. For example, the \acrfull{wrf} \cite{MesoscaleMicroscaleMeteorologyLaboratoryWeatherModel} model is used to predict the rainfall, temperature, and wind direction. Typically multiple \acrshort{nwms} are executed on the same data with different configuration parameters. Then using the real-time observations, the most fitting model is selected. After removing the bias, the data extracted from a \acrshort{nwm} are then fed into another set of models/simulators. For example, FLO2D, a hydrologic model that predicts surface water level, could use precipitation estimates from \acrshort{wrf} to forecast flood levels.

\acrfull{curw} \cite{CUrWSL2017Lanka} is tasked with the mission to reduce water-related natural disasters in Sri Lanka. 
However, as the existing weather data integration systems are not open source and not flexible enough to change without much help from the maintainers, the \acrshort{curw} \cite{CUrWSLObservedSL} is implementing own system with the idea of long-term maintainability.
%\acrshort{curw} \cite{CUrWSLObservedSL} is implementing own system with the idea of long-term maintainability rather than depending on existing weather data integration systems. \gkc{Because of existing weather systems are not open source to configure as required, and not flexible enough to change without much help from the maintainers.} 
\acrshort{curw} collects real-time data from weather stations and other relevant entities and generates daily forecasts. Then the generated results are shared with relevant public departments responsible for disaster planning, risks mitigation, and disaster management. One of the future goals of \acrshort{curw} is to open up data to the public enabling both societal and commercial use cases.

@@

\acrfull{dias} \cite{Kawasaki2018DataReduction} and \acrfull{madis} \cite{Macdermaid2005ArchitectureP2.39} are two of the well-known Weather Data Integration and Assimilation (WDIA) systems. However, those are proprietary systems. Further, while \acrfull{fews} \cite{Werner2013TheSystem} is free to use, but it is not open source. While a few other data integration systems exists, most of them are also proprietary or closed source. Thus, it is difficult and costly to customize such systems for an island like Sri Lanka which experience different kinds of weather seasons over the year \cite{NaveendrakumarFiveLanka}.
For example, \acrshort{fews} does not support FLO2D model integration, which is used by \acrshort{curw} to forecast water level.
To address such issues, there is also an initiative to implement an open-data open-model framework called Meta Scientific Modeling (MSM). 
But the existing WDIA systems and emerging systems are also based on the monolithic or client service architecture; hence, cannot benefit from technologies such as cloud computing to achieve high scalability and availability. Moreover, such systems cannot be updated or reconfigured without taking the system off line. Furthermore, these systems are platform specific. For example, while \acrshort{fews} can run on either Windows or Linux, downtime is needed to introduce any changes to the system.

### 2021-01-01 lit/sec_1-fews.tex
After Dr. Dilum reviewed the Viva thesis

\cref{fi:fews_schematic} shows the connection between a forecasting system to real-time data integration systems and dissemination systems. \gkc{The \acrshort{fews} is capable of integrate data from different sources via import modules into the central servers, and allow forecasters to forecasting and disseminate the results into the dissemination systems.} One of the widely practiced use cases is to use meteorological forecast data to estimate precipitation, and then use a hydrological and hydraulic model chain to predict the surface water level affected. The hydrological and hydrodynamic models should design based on geographical surface data. The ground surface should be examined geographically and divided into catchments. Based on the affected catchment, it can be further divided into sub-catchments to reduce the complexity of the simulation task, and reduce simulation time with parallel processing. Ideally, the forecasting system flexible enough to allow modification of models and data while keeping the system interaction method constant as working with multiple forecasters.

>>>>

\cref{fi:fews_schematic} shows the connection between a forecasting system to real-time data integration systems and dissemination systems. The \acrshort{fews} can integrate data from different sources via import modules into the central servers, and allow forecasters to forecast and disseminate the results into the dissemination systems. One of the widely practiced use cases is to use meteorological forecast data to estimate precipitation, and then use a hydrological and hydraulic model chain to predict the surface water level affected. The hydrological and hydrodynamic models should design based on geographical surface data. The ground surface should be examined geographically and divided into catchments. Based on the affected catchment, it can be further divided into sub-catchments to reduce the complexity of the simulation task, and reduce simulation time with parallel processing. Ideally, the forecasting system flexible enough to allow modification of models and data while keeping the system interaction method constant as working with multiple forecasters.

### 2021-01-01 lit/sec_3-other-systems.tex
After Dr. Dilum reviewed the Viva thesis

\acrfull{msm} framework develops an open system architecture which consists of open-data concept to access heterogeneous data sources via data agents, and open-model concept to allow individual model handling via model agents. Even through these features are provided by other systems such as \acrshort{fews}, they do not provide decentralized model execution and data integration. Specifically, this framework provides a graphical user interface for workflow creation similar to \acrshort{lead}, rather than providing a configuration-based workflow system as in \acrshort{fews}. In our research, we are not mainly focusing on implementing a workflow engine, we will not look deep into such features here.
Then \acrshort{msm} tries to effectively address the drawbacks of the existing modeling systems by providing features such as an independent open-source and graphical-based workflow engine, facilitating  users to obtain data directly from popular sources, decentralized model administration by modelers, and enables models to be freely connected with each other in the \acrshort{msm} framework.

@@

\cref{fi:msm_architecture} shows the \acrshort{msm} system framework from a top-level design point of view, and it is mainly composed of the \acrshort{msm} core which is an interface with the workflow engine, data agents, and model agents. Thus, the \acrshort{msm} core can become the bottleneck of the system, and a single point of failure in critical situations. The \acrshort{msm} core also controls all other components and by using corresponding model agents and data agents allows it to connect with external models and remote data sources. The \acrshort{msm} core is connected to the workflow engine and provides the end-user with all workflow control capabilities, and it contains data persistence services that can be used by all \acrshort{msm} components. Another drawback is, implementing new data agents and model agents are programming language-dependent. In order to implement Model Agents and Data agents, users need to inherit \acrshort{msm} classes in the provided language. Then users need to upload the code into a folder (not a containerized mechanism, or continuous service deployments).

The open-data architecture of \acrshort{msm} is using to access multiple and heterogeneous external data sources. The users are able to integrate data from any source via data agents, but the data agents are depending on inheriting a generic class to do so. Which means there is a language dependency. The open model architecture allows users to independently develop and deploy their own model via model agents. Here also users have to use a generic agent provided by the \acrshort{msm} framework, but the models can be implemented independently. Users can deploy new model agents by copy pasting the source code to a specific folder, but it is possible to run the models which can be written in C, C++, FORTRAN, Python, Java, etc. During the model execution, it follows a flow similar to the Delft-FEW model adapter with a prepared input file, execute the model, and processing output files. The \acrshort{msm} framework has a data management layer, which stores the data as an in memory and local data caching, featuring a non-relational database scheme that supports high throughput. It stores particular timeseries data as msmDataSets, and creates a unique identifier for each msmDataSet. 
The msmDataSets storage framework contains the time-steps. Each time-step contains a reference to the msmDataSet it belongs to, and a Document (the data oriented structure) with a 2D image.

In this paper, it presents a use case of using the \acrshort{msm} system, and does not provide detailed evaluation regarding a performance load testing. In this use case, it demonstrates the usefulness and effectiveness of the framework in various aspects of scientific modeling activities. The msm has only been fully tested on Windows 10 operating system, which means the framework has some operating system dependency. Further, It is not very clear how is it parallel running data agents, model agents, and the \acrshort{msm} core are. Even though it mentioned it follows the decentralized agents approach, based on the available data we can conclude it also has a monolithic distributed architecture.

>>>>

\acrfull{msm} framework developed an open system architecture that consists of open-data concept to access heterogeneous data sources via data agents, and open-model concept to allow management of individual model via model agents. Even though other systems like \acrshort{fews} provide these features, they do not provide decentralized model execution and data integration. Specifically, this framework provides a graphical user interface for workflow creation similar to \acrshort{lead}, rather than providing a configuration-based workflow system as in \acrshort{fews}. \acrshort{msm} tries to effectively address the drawbacks of the existing modeling systems by providing features such as an independent open-source and graphical-based workflow engine, facilitating  users to obtain data directly from popular sources, decentralized model administration by modelers, and enables models to be freely connected with each other in the \acrshort{msm} framework. However, in our research, we focus on data ingress and egress aspects than the workflow engine.

\cref{fi:msm_architecture} shows the \acrshort{msm} framework from a top-level design point of view, and it is mainly composed of the \acrshort{msm} core which is an interface with the workflow engine, data agents, and model agents. Thus, the \acrshort{msm} core can lead to a bottleneck and a single point of failure in critical situations. The \acrshort{msm} core also controls all other components, and by using corresponding model agents and data agents allows it to connect with external models and remote data sources. The \acrshort{msm} core is connected to the workflow engine and provides the end-user with all workflow control capabilities, and it contains data persistence services that can be used by all \acrshort{msm} components. Another drawback is, implementing new data agents and model agents are programming language dependent. To implement model and data agents, users need to inherit \acrshort{msm} classes in the provided language. Then the users need to upload the code into a folder (not a containerized mechanism, or continuous service deployments).

@@

The open-data architecture of \acrshort{msm} is used to access multiple and heterogeneous external data sources. The users are able to integrate data from any source via data agents, but the data agents are depending on inheriting a generic class to do so. Which means there is a language dependency. The open model architecture allows users to develop and deploy their own model via model agents. Here users also have to use a generic agent provided by the \acrshort{msm} framework, but the models can be implemented independently. Users can deploy new model agents by copy-pasting the source code to a specific folder, but it is possible to run the models written in languages such as C, C++, FORTRAN, Python, and Java. The model execution follows a flow similar to the \acrshort{fews} model adapter with a prepared input file, execute the model, and processing output files. The \acrshort{msm} framework has a data management layer that stores the data as an in memory and local data caching, featuring a non-relational database scheme that supports high throughput. It stores particular timeseries data as msmDataSets, and creates a unique identifier for each msmDataSet. The msmDataSets storage framework contains the time-steps. Each time-step contains a reference to the msmDataSet it belongs to, and a document (the data-oriented structure) with a 2D image.

While the authors presented a use case of using the \acrshort{msm} system, detailed evaluation regarding its performance was not given. Nevertheless, it demonstrates the usefulness and effectiveness of the framework in various aspects of scientific modeling activities. The \acrshort{msm} has only been fully tested on Windows 10 operating system, which means the framework has some operating system dependency. Further, details on how it parallellize data agents, model agents, and the \acrshort{msm} core are not presented. Even though the authors mentioned that \acrshort{msm} follows a decentralized agents approach, it seems to have a monolithic distributed architecture based on the available data.

### 2021-01-01 method/sec_0_introduction.tex
After Dr. Dilum reviewed the Viva thesis

This Chapter presents the design decisions taken while implementing \acrfull{wdias} and how we adopt from the systems analyzed in \cref{ch:literature}.
\cref{se:high_level_design} presents a brief discussion on how  the weather forecasting is happening and why does it need to interact with multiple data formats.
\cref{se:architectural_decisions} shows how the architectural design of the proposed solution evolved while designing and implementing \acrshort{wdias}.
\cref{se:microservice} presents microservice architectural concepts followed while developing the \acrshort{wdias}. Then the reasons behind developing \acrshort{wdias} database structure and technologies used are discussed in  \cref{se:db_struct}.
\cref{se:data_preprocess} discusses how \acrshort{wdias} enables data preprocessing and adding capabilities using extensions. Finally, \cref{se:query} includes details on how to perform timeseries search and geo-based queries on the \acrshort{wdias}.

>>>>

This Chapter presents the design decisions taken while implementing \acrfull{wdias} and how we adopt from the systems analyzed in \cref{ch:literature}.
\cref{se:high_level_design} presents a brief discussion on how  the weather forecasting is happening and why does it need to interact with multiple data formats.
\cref{se:architectural_decisions} shows how the architectural design of the proposed solution evolved while designing and implementing \acrshort{wdias}.
\cref{se:microservice} presents microservice architectural concepts followed while developing \acrshort{wdias}. Then the reasons behind developing \acrshort{wdias} database structure and technologies used are discussed in  \cref{se:db_struct}.
\cref{se:data_preprocess} discusses how \acrshort{wdias} enables data preprocessing and adding capabilities using extensions. Finally, \cref{se:query} includes details on how to perform timeseries search and geo-based queries on \acrshort{wdias}.

### 2021-01-01 method/sec_1-architectural_decisions.tex
After Dr. Dilum reviewed the Viva thesis

The \acrshort{wdias} used \acrfull{k8s} as the container orchestration system. \acrshort{k8s} is an open-source tool for automating deployment, scaling, and management of containerized applications [15]. It groups the containers that make up an application into logical units for easy management and detection. Using \acrshort{k8s}, we deploy each microservice as a container and manage it by providing scalability and availability.

@@

>>>>

\acrshort{wdias} used \acrfull{k8s} as the container orchestration system. \acrshort{k8s} is an open-source tool for automating deployment, scaling, and management of containerized applications [15]. It groups the containers that make up an application into logical units for easy management and detection. Using \acrshort{k8s}, we deploy each microservice as a container and manage it by providing scalability and availability.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Architecture Comparison with Existing Systems

\cref{fi:architecture_comparison} shows a comparison between existing systems and WDIAS concerning software architectural aspects. Here, we compare systems under categories such as data storage, performance, architecture, and other features. The \acrshort{fews} and \acrshort{msm} frameworks store the data on a central data service, which could become a bottleneck while handling large volumes of data in a critical situation. Other systems can use a pool of resources allowing them to increase the performance by adding more resources to the system. Regarding the performance of the systems, we consider scalability, availability, cloud computing support, and system deployment capabilities. All the systems can scale up to some extent by adding resources beforehand or in real-time. However, LEAD, DIAS, and WDIAS could scale on-demand by adding more resources and support automatic elasticity (i.e., scaling up and down). LEAD, MADIS, DIAS, and WDIAS have high availability, as they use a resource pool that can run on different locations. Further, these systems are capable of automatically handling partial system failures enabling fault tolerance. When we consider cloud computing support, the system should support easy integration with clouds. It is possible to rent virtual servers and deploy the local services in the cloud infrastructure, but some systems may not allow access to them remotely. Even though some systems can deploy in such a manner, they may not be capable of automatically allocating resources on demand. Hence, we cannot consider that those systems can utilize cloud infrastructures fully. Furthermore, we can easily maintain a system if it consists of independent subsystems, and each subsystem is capable of deploying or upgrading independently. Independent service deployment is a more flexible and better way to handle this, rather than copy-pasting codes to deploy changes.

\begin{figure}[htp]
    \centering
    \includegraphics[width=1.05\textwidth]{method/misc/architecture_comparison_pros_cons.pdf}
    \caption{Comparison of features among existing weather data assimilation and integration systems and \acrshort{wdias}. %An "X" in the table represents a feature offered by the corresponding tool. An "-" in the table represents that information available are not enough to make an statement, and empty presents a feature is not offered. 
    }
    \label{fi:architecture_comparison}
\end{figure}

We discuss the architecture of the systems in terms of their design, dependencies of the subsystems, extensibility, availability of source code, and data preprocessing capabilities specific to the weather domain. While other existing systems use distributed client-server architecture or monolithic distributed architecture, LEAD uses the SOA architecture. However, WDIAS has the advantage of using modern cloud computing technologies to implement microservices as containerized applications. The architecture implementation of independent subsystems allows users to maintain subsystems independently. Further, WDIAS enables different subsystems to be implemented using different technology stacks. Specially WDIAS and MSM provide an open-source framework that gives users the capability to enhance or modify the system as per the requirement. These systems provide data preprocessing capabilities for weather data and the ability to add new preprocessing capabilities. However, WDIAS is also capable of providing auto-scaling with data preprocessing. Workflow engine and model execution support are some other extra features that are helpful with weather forecasting. However, the WDIAS focuses on implementing a WDIA system with supporting main features, and it provides an extendable open framework to integrate such features easily.

### 2021-01-01 method/sec_3-microservice.tex
After Dr. Dilum reviewed the Viva thesis

Figures \ref{fi:wdias_micro_on_demand} and \ref{fi:wdias_micro_async} show the system interaction of the WDIA microservices architecture while handling requests. Each circle in the figures represents a microservice in the \acrshort{wdias}, and each microservice is implemented as a containerized application. The left side of \cref{fi:wdias_micro_on_demand} shows the import modules of the \acrshort{wdias}, and the right side shows the export modules. Also, we implemented each module as a microservice to independent maintainability and upgradability. We followed the concept of smart endpoints and dumb pipes explained in \cref{subse:dumb_pipes}, and each import microservice only does the specific task of converting and forwarding the request to the correct data adapter module. For scalar, vector, and grid data types, we created adapter microservices that are optimized for storing such kinds of data. The metadata data of the timeseries store using an RDBMS, which gives more performance over retrieving timeseries metadata. The metadata is cached using an in-memory database to fast access while implementing the database structure. The system generates a unique identifier for each timeseries, and throughout the \acrshort{wdias} system, all microservices use it to handle data for fast access.

@@

\cref{fi:wdias_micro_separation} shows the clear separation of microservices into the modules of the \acrshort{wdias}. Each adapter has a separate database, and the database is hosted separately for the high performance and to avoid interference.
Apart from that, as we further discuss in \cref{se:data_preprocess}, the extension modules are running independently, such as interpolation, transformation, and validation, and we will also discuss in \cref{se:data_preprocess}. The extension adapter allows users to register new event-based or time-based triggers for the extensions. The extension scheduler triggers the events based on time and the extension handler triggers events based on data change.

>>>>

Figures \ref{fi:wdias_micro_on_demand} and \ref{fi:wdias_micro_async} show the system interaction of the WDIA microservices architecture while handling requests. Each circle in the figures represents a microservice in \acrshort{wdias}, and each microservice is implemented as a containerized application. The left side of \cref{fi:wdias_micro_on_demand} shows the import modules of \acrshort{wdias}, and the right side shows the export modules. Also, we implemented each module as a microservice to independent maintainability and upgradability. We followed the concept of smart endpoints and dumb pipes explained in \cref{subse:dumb_pipes}, and each import microservice only does the specific task of converting and forwarding the request to the correct data adapter module. For scalar, vector, and grid data types, we created adapter microservices that are optimized for storing such kinds of data. The metadata data of the timeseries store using an RDBMS, which gives more performance over retrieving timeseries metadata. The metadata is cached using an in-memory database to fast access while implementing the database structure. The system generates a unique identifier for each timeseries, and throughout the \acrshort{wdias} system, all microservices use it to handle data for fast access.

@@

\cref{fi:wdias_micro_separation} shows the clear separation of microservices into the modules of \acrshort{wdias}. Each adapter has a separate database, and the database is hosted separately for the high performance and to avoid interference.
Apart from that, as we further discuss in \cref{se:data_preprocess}, the extension modules are running independently, such as interpolation, transformation, and validation, and we will also discuss in \cref{se:data_preprocess}. The extension adapter allows users to register new event-based or time-based triggers for the extensions. The extension scheduler triggers the events based on time and the extension handler triggers events based on data change.

### 2021-01-01 method/sec_4-db_struct.tex
After Dr. Dilum reviewed the Viva thesis

The architecture and the storage management of the \acrshort{wdias} is designed and developed based on the weather data. We need to look into the weather data timeseries deeply to understand the underline design decisions of the \acrshort{wdias} database structure.

@@

Based on the above interpretation, a timeseries have data points ordered in time. Other than that, timeseries have additional information that is known as \emph{timeseries metadata}, which can use to identify one timeseries from another uniquely. As an example, we can place a weather station on a \emph{location}, and it can measure different types of weather data that are called \emph{parameters} such as temperature, precipitation, and wind direction. Each parameter can produce different timeseries. Even we can use the location as metadata to identify the timeseries, but that is not enough to distinctly identify the parameters. Thus, we need to use a set of attributes to identify a timeseries uniquely. Next, we describe key attributes that are used by the \acrshort{wdias} to identify a timeseries uniquely.

@@

\subsubsection{Value Type}
Scalar, vector, grid are the value types that are interested in the \acrshort{wdias}.

\subsubsection{Location}
Location of the timeseries. All locations have a unique String identifier called \emph{locationId}. Further, two types of location types are interested in the \acrshort{wdias}, namely:
\begin{itemize}
  \item \emph{Point locations} -- Contains a name that is human readable. latitude(lat) and longitude(lon) of the location on the earth surface. Following is an example of define a location using \acrfull{json}.
\begin{lstlisting}[language=Python]
{
    "locationId": "wdias-hanwella",
    "name": "Hanwella",
    "lat": 6.909722222,
    "lon": 80.08166667
}
\end{lstlisting}
  \item \emph{Regular Grid locations} -- Contains a name that is human readable. The grid presents by dividing into equal size cells. Thus, it needs the number of rows and columns. 
  And the location of the first cell and the width and height of it. Below \acrshort{json} format shows an example of the regular grid location. 
\begin{lstlisting}[language=Python]
{
    "locationId": "wdias_kelani_basin",
    "description": "Kelani Basin",
    "rows": 120,
    "columns": 139,
    "geoDatum": "Kandawala",
    "gridFirstCell": {
        "firstCellCenter": {
            "x": 397074.0,
            "y": 504875.0
        },
        "xCellSize": 250.0,
        "yCellSize": 250.0
    }
}
\end{lstlisting}
  \item \emph{Irregular Grid locations} -- (endpoints are available within the WDIAS system, but skipped the implementation since it is out of the interest of the scope.)
\end{itemize}

@@

\cref{fi:database_structure} shows the \acrshort{wdias} database structure.
The following sections describe the database structure of \acrshort{wdias} with compare to the design with more details on how is support the microservice architecture.

@@

Two individual instances of InfluxDB used for scalar adapter and vector adapter services in the \acrshort{wdias} system. For the scalability, the system uses \emph{database per service}, as mentioned in \cref{subse:database_per_service}. If users want further performance, the database can split into another set of partitions based on other key attributes based on the z-axis scaling described in \cref{subse:scale_cube}. As an example, scalar and vector adapters can further split into more services based on other key attributes such as timeseries type.

@@

We used MongoDB geospatial features to support geo-queries in the \acrshort{wdias}. Also, we indexed the timeseries metadata with other indexing features to serve external timeseries metadata queries. If particular timeseries metadata not present in the query adapter, then fetch metadata from the metadata adapter and indexed in the MongoDB. Vise a versa; when a new timeseries created in the metadata adapter, it triggers an event to index metadata in the query adapter.
Here we used the \emph{y-axis scaling} concept by splitting metadata search into multiple services. Within the system, it uses in-memory database cache to support internal metadata access, and external search queries fulfill by the query adapter with using MongoDB.

>>>>

The architecture and the storage management of \acrshort{wdias} is designed and developed based on the weather data. We need to look into the weather data timeseries deeply to understand the underline design decisions of the \acrshort{wdias} database structure.

@@

Based on the above interpretation, a timeseries have data points ordered in time. Other than that, timeseries have additional information that is known as \emph{timeseries metadata}, which can use to identify one timeseries from another uniquely. As an example, we can place a weather station on a \emph{location}, and it can measure different types of weather data that are called \emph{parameters} such as temperature, precipitation, and wind direction. Each parameter can produce different timeseries. Even we can use the location as metadata to identify the timeseries, but that is not enough to distinctly identify the parameters. Thus, we need to use a set of attributes to identify a timeseries uniquely. Next, we describe key attributes that are used by \acrshort{wdias} to identify a timeseries uniquely.

@@

\subsubsection{Value Type}
Scalar, vector, grid are the value types that are interested in \acrshort{wdias}.

\subsubsection{Location}
Location of the timeseries. All locations have a unique String identifier called \emph{locationId}. Further, two types of location types are interested in \acrshort{wdias}, namely:
\begin{itemize}
  \item \emph{Point locations} -- Contains a name that is human readable. latitude(lat) and longitude(lon) of the location on the earth surface. Following is an example of define a location using \acrfull{json}.
\begin{lstlisting}[language=Python]
{
    "locationId": "wdias-hanwella",
    "name": "Hanwella",
    "lat": 6.909722222,
    "lon": 80.08166667
}
\end{lstlisting}
  \item \emph{Regular Grid locations} -- Contains a name that is human readable. The grid presents by dividing into equal size cells. Thus, it needs the number of rows and columns. 
  And the location of the first cell and the width and height of it. Below \acrshort{json} format shows an example of the regular grid location. 
\begin{lstlisting}[language=Python]
{
    "locationId": "wdias_kelani_basin",
    "description": "Kelani Basin",
    "rows": 120,
    "columns": 139,
    "geoDatum": "Kandawala",
    "gridFirstCell": {
        "firstCellCenter": {
            "x": 397074.0,
            "y": 504875.0
        },
        "xCellSize": 250.0,
        "yCellSize": 250.0
    }
}
\end{lstlisting}
  \item \emph{Irregular Grid locations} -- (endpoints are available within the WDIAS system, but skipped the implementation since it is out of the interest of the scope.)
\end{itemize}

@@

\cref{fi:database_structure} shows \acrshort{wdias}'s database structure.
The following sections describe the database structure of \acrshort{wdias} with compare to the design with more details on how is support the microservice architecture.

@@

Two individual instances of InfluxDB used for scalar adapter and vector adapter services in \acrshort{wdias} system. For the scalability, the system uses \emph{database per service}, as mentioned in \cref{subse:database_per_service}. If users want further performance, the database can split into another set of partitions based on other key attributes based on the z-axis scaling described in \cref{subse:scale_cube}. As an example, scalar and vector adapters can further split into more services based on other key attributes such as timeseries type.

@@

We used MongoDB geospatial features to support geo-queries in \acrshort{wdias}. Also, we indexed the timeseries metadata with other indexing features to serve external timeseries metadata queries. If particular timeseries metadata not present in the query adapter, then fetch metadata from the metadata adapter and indexed in the MongoDB. Vise a versa; when a new timeseries created in the metadata adapter, it triggers an event to index metadata in the query adapter.
Here we used the \emph{y-axis scaling} concept by splitting metadata search into multiple services. Within the system, it uses in-memory database cache to support internal metadata access, and external search queries fulfill by the query adapter with using MongoDB.
